Reachability analysis for continuous one-counter automata

### 2021-05-21 Documentation/paper.tex
finished automaton section

@@

>>>>

\usepackage{subfig}

@@

\subsection{Automaton input}
For the approach described within this paper, an input of an automaton using the dot language part of Graphviz was assumed \cite{10.1007/3-540-45848-4_57}. This language is a very versatile and easily understandable format in which automata can be defined. It was also used as the output language in the preceding research which automatically made it the most interesting format to support. No other formats were considered as the conversion from a textual representation to an instance of the \textit{Automaton} class is not considered to be the core of the research. The logic used for the dot language is furthermore identical to the logic that would need to be applied to read any other automaton specification, with the only difference lying in the language specific aspects of the specification.

\subsubsection{Edge transformation}
Edges are represented within the dot language by a single line per edge using the format specified below. The \textit{origin} and \textit{end} tokens specify the start and end node of the transition. The square brackets encapsulate optional arguments, of which the label and xlabel arguments matter most, as either can be used to specify the label of the transition, and thereby the operation that we want to have attached to the edge.

\begin{gather*}
	\textless origin \textgreater \rightarrow \textless end\textgreater [label=\textless label\textgreater, xlabel=\textless label\textgreater, ...]
\end{gather*}

This configuration can easily be converted into edge instances, but the labels can still represent several things. To allow any possible automaton specification, we will simply filter the labels using a regex for operations and labels. If it does not match, we assume the label to be an actual name label, if it does match, the label is perceived as a mathematical formula and it will be registered as such.

To closely match the specification of the theoretical proof on which this approach is based \cite{blondin2021continuous}, we require edges to only contain operations and not conditions. In case an edge is encountered with a condition as a label, this will be resolved by introducing a new node, \textit{inbetween}, to which this condition will be added. This node will be added as an intermediary step and the condition label will be attached to this node. By ensuring that this node will only have one incoming edge and one outgoing edge we preserve the fact that the condition must hold in case one desires to from from the \textit{origin} node to the \textit{end} node.

Below is an example of an edge which does not correspond to the assumptions made within the original paper as the edge between \textit{q0} and \textit{q1} has a conditional label. This is resolved in the image on the right by introducing \textit{\_0} without losing the original meaning of the edge in the first image.

\begin{figure}[h]%
	\centering
	\subfloat{
		\begin{tikzpicture}[auto, >=latex, node distance = 1 cm]
		\tikzstyle{round} = [thick, draw=black, circle, font=\small]
		\tikzstyle{invis} = [draw=none, font=\small]

		\node[round] 	at (0, 0) 		(q0) 	{$q0$};
		\node[round] 	at (3, 0)		(q1) 	{$q1$};

		\path[->]
		(q0)	edge	node {$>= 0$}	(q1);
		\end{tikzpicture}
	}
	\qquad
	\subfloat{
		\begin{tikzpicture}[auto, >=latex, node distance = 1 cm]
		\tikzstyle{round} = [thick, draw=black, circle, font=\small]
		\tikzstyle{invis} = [draw=none, font=\small]

		\node[round] 	at (0, 0) 		(q0) 	{$q0$};
		\node[round] 	at (3, 0)		(q1) 	{$q1$};
		\node[invis]	at (2.3, 2.5)		(q3)	{$>=0$};
		\node[round] 	at (1.5, 2) 	(q4) 	{$\_0$};

		\path[->]
		(q0)	edge	node {$ $}	(q4)
		(q4)	edge	node {$ $}	(q1);
		\end{tikzpicture}
	}
\end{figure}

\subsubsection{Node transformation}
Nodes do not require to be specified using a separate node specification rule, and will simply be generated in case they are used within an edge. Node specifications function specifically for the cases in which additional attributes need to be added to nodes. To support operations as well as general labels, the only two attributes which will be tracked from node specification are the \textit{label} and \textit{xlabel} arguments. These labels will be matched with regexes specifying the supported formats for operations and conditions. In case the label does not match either, it is still tracked as regular node labels are still relevant for reachability reporting. For the approach discussed within this paper the node labels represent the code line from which they were generated and are therefore needed to analyse code line reachability. All node specifications will follow the format given below where the square bracket can contain a number of optional arguments.

\begin{gather*}
\textless node \textgreater [label=\textless label\textgreater, xlabel=\textless label\textgreater, ...]
\end{gather*}

To match the specification of the theoretical approach specified by Michael Blondin et Al. \cite{blondin2021continuous}, nodes can only contain conditions and not operations. In case a node \textit{origin} is found that has an operation as its label we introduce a new node \textit{inbetween}. By connecting a new edge from the \textit{origin} node to the \textit{inbetween} node we can add the label that was originally attached to \textit{origin} to the new edge. To ensure that the automaton does not lose any meaning we will then move every edge that would normally go out from \textit{origin} to \textit{inbetween}.

In the image below, an example can be seen where node \textit{q1} is not conformant to the original specification. This is resolved by introducing \textit{\_0} from which the edge that was originally going from \textit{q1} to \textit{q2} gets reattached.

\begin{figure}[h]%
	\centering
	\subfloat{
		\begin{tikzpicture}[auto, >=latex, node distance = 1 cm]
		\tikzstyle{round} = [thick, draw=black, circle, font=\small]
		\tikzstyle{invis} = [draw=none, font=\small]

		\node[round] 	at (0, 0) 		(q0) 	{$q0$};
		\node[round] 	at (2, 0)		(q1) 	{$q1$};
		\node[invis]	at (2.5, 0.5)	(q3)	{$+1$};
		\node[round] 	at (4, 0) 		(q2) 	{$q2$};

		\path[->]
		(q0)	edge	node {$ $}	(q1)
		(q1)	edge	node {$ $}	(q2);
		\end{tikzpicture}
	}
	\qquad
	\subfloat{
		\begin{tikzpicture}[auto, >=latex, node distance = 1 cm]
		\tikzstyle{round} = [thick, draw=black, circle, font=\small]
		\tikzstyle{invis} = [draw=none, font=\small]

		\node[round] 	at (0, 0) 		(q0) 	{$q0$};
		\node[round] 	at (2, 0)		(q1) 	{$q1$};
		\node[round] 	at (4, 0) 		(q2) 	{$q2$};
		\node[round] 	at (2, 2) 		(q4) 	{$\_0$};

		\path[->]
		(q0)	edge	node {$ $}	(q1)
		(q4)	edge	node {$ $}	(q2)
		(q1)	edge	node {$+1$}	(q4);
		\end{tikzpicture}
	}
\end{figure}

\subsubsection{Initial node}
For reachability analysis to be relevant there needs to be an initial node. In automata, the initial node is often represented by a transition going from nowhere towards a single node. This is however not a built in feature within the dot language. One of the solutions to overcome this is to create a node and make it invisible, an edge can then be drawn from this invisible node to the actual initial node. Other approaches are potentially possible but were not explored. Any alternative implementations could easily be introduced instead of the considered format without interchanging any of the other components.

As a consequence of this feature invisible nodes are assumed to exist within the automaton but these are irrelevant to our analysis. Invisible nodes are by definition not relevant for reachability analysis and are therefore not considered as actual nodes.

\subsection{Continuous one counter automata reachability}
In this section we will give a description of the solution that was used to determine the reachability for continuous one counter automata. This will be done by showing the approach that was taken, as well as an application to show its usage. This approach is a practical implementation of the theoretical approach described by Michael Blondin et al. \cite{blondin2021continuous}.

\begin{landscape}
	\begin{figure}
		\centering
		\begin{tikzpicture}[auto, >=latex, font=\footnotesize\scshape]
			\tikzstyle{center} = [draw=black, rectangle, text width=5.8em]
			\tikzstyle{arrow} = [thick,->,>=stealth]

			\node[center]	at (0, 0) 		(q0) 	{Intersect \\input: x,y \\output: z};

			\node[center]	at (3, 6) 		(q1) 	{x is empty};
			\node[center]	at (3, 5)		(q2)	{y is empty};
			\node[center]	at (3, 2)		(q5)	{$x_\bot == 2$};
			\node[center]	at (3, 2)		(q5)	{$x_\bot == 2$};

			\node[center]	at (7, 6)		(q3)	{$z == [0, 0, 1, 1]$};
			\node[center]	at (7, 5)		(q4)	{$z == [0, 0, 1, 1]$};
			\node[center]	at (7, 4)		(q6)	{$y_\bot == 2$};
			\node[center] 	at (7, 3)		(q8)	{$y_b <= x_t$};
			\node[center] 	at (7, 1)		(q11)	{$y_b > x_t$};
			\node[center]	at (7, 0)		(q13)	{$x_\top == 2$};

			\node[center]	at (11, 4)		(q7)	{$z_\bot == 2$};
			\node[center] 	at (11, 3)		(q9)	{$z_b == y_b$};
			\node[center] 	at (11, 2)		(q10)	{$z_\bot == x_\bot$};
			\node[center]	at (11, 1)		(q12)	{$z == [0, 0, 1, 1]$};
			\node[center]	at (11, 0)		(q14)	{$z == y$};

			\draw [arrow] (q0) |- (q1);
			\draw [arrow] (q0) |- (q2);
			\draw [arrow] (q0) |- (q5);
			\draw [arrow] (q1) |- (q3);
			\draw [arrow] (q2) |- (q4);
			\draw [arrow] (q5) |- (q6);
			\draw [arrow] (q6) |- (q7);
			\draw [arrow] (q5) |- (q8);
			\draw [arrow] (q8) |- (q9);
			\draw [arrow] (q8) |- (q10);
			\draw [arrow] (q5) |- (q11);
			\draw [arrow] (q11) |- (q12);
			\draw [arrow] (q5) |- (q13);
			\draw [arrow] (q13) |- (q14);
		\end{tikzpicture}
	\end{figure}
\end{landscape}
